How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-28 tse-paper/content.tex
improve lit discussion and emp comparison study explanation

%\section{introduction}
% TODO: What about the alternate solution, installing plugins to report
% precise data? Ie., maven plugins
% - Lots of effort
% - not a generic solution
% - how to analyze history where such plugins were not installed
% - unclear which information you want

% TODO It would be futile to compare our techniques with one of the
% existing manual regex parsers, as (a) as our literature survey showed,
% their scope is mainly limited to the Java ecosystem and (b) whether
% they would work or not would be randomly based on whether the project
% we select exhibits features their regular expressions were trained for.
% Finally, they are not really automatic

@@

In the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that automatic software build
processes create.
As a treasure trove of information~\cite{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis tools.
The
contents and format of build logs can vary highly depending on which tools
are involved in the build process, how they are
configured, and as projects and build tools evolve over
time~\cite{staahl2014modeling}.
Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

When a build or one of its steps fails, developers typically scavenge
the logs for information related to the source of the error.
This
manual activity is tedious and prone to
error~\cite{santolucito2018statically}.
As a single build log can be
over 50 megabytes large~\cite{beller2017oops}, finding the small chunk
of information in it which is linked to the actual error, is akin to
searching for a needle in a haystack.
The large numbers of unrelated, but ambiguous warnings that many build
logs regularly contain, further compounds this problem.
In addition to helping
understand and fix build failures, the information richness of build
logs enables a plethora of other applications: with the help of build
logs, we can better understand and categorize CI failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to do ``predictive CI'' on the build or test outcome
and
duration~\cite{ni2017cost,bisong2017built,haghighatkhah2018test,
machalica2019predictive},
or investigate the role of automated static analysis tools (ASATs,
\cite{beller2016analyzing}) in CI
builds~\cite{zampetti2017open}.

@@

Despite its central role in helping developers and in enabling
novel, on-ward processing applications,
the process of how to
automatically analyze build logs has thus far not been systematically
overviewed as a research topic, bringing us to the first
research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What is the state-of-the-research on analyzing build logs?
\end{simplebox}

To answer this question, we start our investigation---pictured in
\Cref{fig:overview}---with an
extensive literature study (step 1).
The literature
survey identifies various strategies to
analyze build logs, among which are developing
custom parsers, using regular expressions, doing manual inspections,
searching for keywords, or performing natural language processing.
However, there is currently no guidance on when to use which of these
approaches.
Unfortunately, our survey also shows that
less than half of the 61 works which rely on build log analysis
describe it in sufficient detail and that
the few available implementations are all based on either a custom parser
or regular expressions.
These two approaches in particular are problematic, as their creation
and maintenance is manual and expensive, does not generalize across build
logs from different build environments, and is, as the experience of
TravisTorrent has shown, prone to error.
In fact, some even ``strongly discourage'' the creation of such
bespoke parsers~\cite{urli2018design}.
In summary, the outcome of our literature survey shows the need for a new
generation of automated techniques to analyze build logs, which are
documented in detail and which are openly available, together with
guidance on when to use which tool.

@@

% M: maybe convert this into a proper ordered list?
% TODO Moritz: not sure what you mean here ?
% I agree that the non-itemized looks better than before
To fill this gap, we adapted and implemented three novel, prototypical
implementations to
automatically analyze build logs.
The ideas behind the implementations are based on promising approaches
adopted from other research fields and automatic build log analysis
techniques:
a technique 1) able to automatically synthesize regular expressions,
called \emph{Program Synthesis by Example (PBE)}, 2) based on information
retrieval, using \emph{Common Text Similarity (CTS)}  and 3)
inspired by the manual adhoc-search for key parts in the log, employing
\emph{Keyword Search (KWS)}.
These three approaches---regular expressions, text similarity,
keyword search---
have different strengths and weaknesses.
As a consequence,
in the second research question we ask:

@@

To answer this question, as a third step,
we conduct an empirical study assessing the performance
of these chunk retrieval techniques in
real-world projects under a variety of conditions.
Our study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses 797 build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages.
Our results show that there is no technique that in general outperforms
the others.
PBE has the best average precision of 95\%, while KWS has the largest
recall with 70\% on average.
However, CTS shows the best F$_{1}$-score
of 51\% and all of the three techniques
clearly outperform a random baseline.

Finally, due to this ambiguity, we develop a set of guidelines
(\Cref{fig:overview}, step 4)
on how to choose the most suitable build log analysis
technique for the task at hand.
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
The results PBE produces are suited for automatic on-ward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a finding---makes it generally
unsuited for automatic on-ward processing.
Instead, KWS requires a human
to further inspect and interpret the output chunk.
A hybrid solution combining the three techniques might be the best
all-round solution.
Finally, we give a roadmap to guide future research in the field
of build log analysis.

In short, this article contributes
\begin{itemize}
\item the first systematic study of the emerging field of build log
analysis.
\item three prototypical open-sourced implementations of
promising chunk retrieval techniques.
\item a large empirical study comparing the three techniques.
\item systematic guidelines on when to use which technique.
\item a detailed research map to inspire future contributions to the
field.
\item a replication package~\cite{brandt2020chunk-replication}.
\end{itemize}

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system
logs which are produced during runtime of the system.
The first difference is that build logs are produced during runtime
of the build of the system, not during runtime of the system itself.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} depicts two
examplary lines from such a system log.

@@

\subsection{Literature Selection}
The first step in a systematic literature study is to
establish which works one wants to cover and how to find
them~\cite{kitchenham2009systematic}.
Since the field of build log analysis is relatively young, we
surveyed as broad a population of scientific material as possible.
We did not limit the sources to specific Software Engineering
venues~\cite{petersen2015guidelines}, but included all sources
monitored by Google Scholar, including Bachelor's theses, Master's
theses, and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases because it is the search engine for scholarly material
with
the broadest and most current index, including preprints.

\Cref{fig:lit-survey} gives an overview of the study selection
process.
An initial Google Scholar search for variations of ``build log''
returned more than 2.2 million results on the 15th of March 2020, many
of which stem
from unrelated fields such as construction or wood working.
We refined the search criteria to exclude such obvious
non-related fields and to exclude works on systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results.
% Our replication package documents the full search
% queries~\cite{brandt2020chunk-replication}.
To make handling so many search results feasible, we employed a
three-pass filtering strategy:

First, we filtered articles based on (a) their title and (b) the
contextual information Google Scholar displayed on its search results
overview page.
We included works at large that seemed to bear a resemblance to software
engineering and building software.
Following Kitchham's protocol~\cite{kitchenham2009systematic}, we excluded
works written in a language unintelligible to the authors
(i.e., not English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

@@

% TODO	everything is in passive in the following
Specifically, we were interested in
\begin{itemize}
  \item what information is retrieved from the logs and
  what the retrieved information is used for (\textbf{RQ1.1})
  \item which techniques are used to process the build logs,
  in how much detail they are described, and whether they are published
  as tools (\textbf{RQ1.2})
  \item what kind of build logs are used and the origin of these
  logs.
(\textbf{RQ1.3})
\end{itemize}

This lead us to the following research questions:
\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What is the state-of-the-research on analyzing build logs?}
% M: All of these are passive
% C: it says resolved, but they still are?
% M: Only the first one still is (don't know how to convert it),
% the others
% are not?
% I exchanged 'extracted' with targeted / analysis
% 'information retrieved' or 'information extracted' will give us
% this is not IR neither IE discussions again
\begin{itemize}[leftmargin=1cm]
  \item[\textbf{RQ1.1:}] What information is targeted in build logs?
  \item[\textbf{RQ1.2:}] Which analysis techniques exist?
  \item[\textbf{RQ1.3:}] Which kinds of build logs are subject to
  analysis?
\end{itemize}
\end{simplebox}

To support the data extraction we created a template that we filled out
for each of the 61 works.
The template contained questions to answer RQ 1.1 through RQ 1.3,
for example ``Is the technique explained in detail?'' and
``What is the source of the build logs?''
As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping study with a
pilot:
both authors extracted data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 56 articles evenly between the two authors
and inspected each articlex as closely as necessary to fill out the
template
with confidence,
starting from occurrences of ``log'' or ``build'' within the article's
text.

@@

\subsection{Descriptive Study Statistics}
To be done.
% TODO moritz: age of papers (graph), kind of
% publication (master, paper)
% published (y, n)

@@

\raisebox{0.8mm}{Information Retrieval} &
\raisebox{0.8mm}{
\cite{hassan2017change,lindqvist2019detection,ren2018automated}
} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\
% TODO: We need to mention in the text describing the Table what we mean
%by ``Analysis''
\raisebox{0.8mm}{Analysis} &
\raisebox{0.8mm}{
\cite{sulir2016quantitative,haghighatkhah2018test,durieux2019critical}
} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\

@@

\Cref{tab:litsur:techniques} presents the techniques that the articles
described for analyzing build logs.
43 of the 61 articles we inspected mentioned how they are analyzing
build logs.
However, only 16 of these described their method in detail.

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''
~\cite{rahman2018impact}.
For such imprecise statements it was often not clear if they describe
a complex parser (including a tokenizer, lexer, etc.) or refer to a
simpler parsing tool based on regular expressions.
If we saw evidence for the latter, we assigned the tag
``regular expression'' instead of the tag ``parser.''
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with manual categorizing build logs and selecting
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.
Another 8\% employed machine learning such as natural language
processing or information retrieval techniques.

@@

Only 7\% of the proposed methods claimed to be language-agnostic, while
several described they are covering multiple source build tools.
The majority of articles (42\%) used logs from a source that no
second article
targeted.
In 38\% of cases, the articles collected build logs from Travis
CI~\cite{travisci2019webpage};
18\% of the articles processed logs from TravisTorrent; and 8\% included
build logs from industrial projects.
The number of logs analyzed varied greatly between the articles
and we were not always able to
extract it confidently.
23 of the articles analyzed more than a thousand build logs, some
investigated up to 122 million logs.
Of the surveyed articles, 31\% claim to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which
Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying},
and several other articles studied.

@@

Our literature survey shows that despite its relatively young age
(first mentioned in a paper in 2001, vast majority after
2016), \textbf{build log analysis is an established technique}
in the literature, with 61 articles making use of it.

% TODO we are extracting failures in the following studies, which is
%mainly what
% literature does, too, right? Highlight here and mention later again

% TODO: this section is crucial.
% maybe we can highlight important findings, by categorizing them either
% as further subsubsections or making it bold in the text?

@@

\textbf{most analysis aims to find out why the build failed}

\textbf{most techniques retrieve chunks from build logs}

\textbf{Most log analysis is automated.} Automated approaches are
necessary as many studies target a large number of builds.
Articles
which explicitly mention a manual analysis of the logs, mostly
restrict their work to few logs or take a ``representative sample to
make manual analysis feasible''~\cite{zolfagharinia2017not}.
\textbf{Developing automated analysis tools requires a lot of effort.}
Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone,'' other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.

We observed \textbf{very little reuse of build log analysis tools.}
This can stem from the
high specialization of the developed tools in regards to supported
% TODO: build environments is used a bunch (also in the following), but
% never defined.
build environments and targeted information.

% TODO: draw a conclusion together with figure 5: if reuse were higher,
%lots of effort could be saved?
Several articles noted the variety in log formats for different build
environments, which requires customizing build log techniques for
every supported build environment.
% The following two sentences are passive, however flipped around
% have a different meaning (we want build env -> study count)
\Cref{fig:litsur:log_producer} shows that only a few distinct
build environments are the target of
many studies and tools (roughly, area A).
The majority of build environments was however only targeted by one
article each (roughly, area B).
From this we follow, that the large tail of
\textbf{seldomly-investigated
environments would benefit from the
existence of a more generic solution.}

% TODO we argue this in the intro!
% problem: we have no data on this
\textbf{techniques vary and seem rather randomly applied}

\textbf{modifying the build output is not helpful}, as well as returning
precise data through other documents than the log.
This requires access to the build configuration, is also specific to
the used build environment and prohibits the analysis of
historical data.

\textbf{TODO embed custom plugin discussion}
% TODO very disconnected here
From this, we follow: 1) there is a niche for custom parsers or even
specialized plugins integrated into the build process to make
post-factum
parsing of logs redundant.
Their advantage is that they can precisely report the information which
is required.
A downside of plugins, which are integrated into the CI flow similar to
Deflaker by Bell et al.~\cite{bell2018deflaker}, is that they are unable
to work on historic
builds, i.e., before the moment they were being deployed and that if
information has not been recorded from the beginning but turns out to
be important, it is impossible to retrieve later.
2) the large tail of seldomly-investigated environments would benefit
from the existence of a more generic
solution.

% We implement build environment agnostic build log analysis
% techniques, which are customized to a specific project by
% the training examples which the user provides.
% In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
% data set we chose for our evaluation covers a broad range of build
% environments.
% % Further, the training examples also specify the log chunk targeted
% by the
% % retrieval, which enable our chunk retrieval techniques to retrieve
% % a broad range of information from build logs.

@@

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
%TODO here we sell our techniques as superior.
% match in intro?
%TODO: really? currently the discussion does not show this 'need'!
The outcome of our literature survey shows the need for a new
generation of techniques to analyze build logs.
In this section we argue how our research sets the first step to
address this need by introducing three techniques that retrieve
specific substrings (chunks) from build logs.
We present defining characteristics of these so called \emph{chunk
retrieval techniques} and present the techniques we developed in
detail.
% TODO At the end of what?
At the end we illustrate how the techniques operate on a concrete
example and sketch out their implementation.

@@

\subsection{The Case for Chunk Retrieval on Build Logs}
In summary, the results of the literature study show the need for
build log
analysis techniques that are \emph{automated},
\emph{agnostic of the build environment}
and can be \emph{configured with little effort}.

>>>>

%\section{introduction}

% Good point, but we don't really need it anywhere:
%
% It would be futile to compare our techniques with one of the
% existing manual regex parsers, as (a) as our literature survey showed,
% their scope is mainly limited to the Java ecosystem and (b) whether
% they would work or not would be randomly based on whether the project
% we select exhibits features their regular expressions were trained for.
% Finally, they are not really automatic

@@

In the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build logs are a textual by-product that these automatic software
builds create.
As a treasure trove of information~\cite{meyer},
build logs contain trace outputs of not only the elementary steps to
``make'' a project---such as retrieving and resolving dependencies and
compiling---but also about ancillary quality assurance steps---such as
testing the software and running automated static analysis tools.
The
contents and format of build logs can vary highly depending on which tools
are involved in the build process, how they are
configured, and as projects and build tools evolve over
time~\cite{staahl2014modeling}.
Typically, though, they are at least a semi-structured and
somewhat stable journal of the commands executed during the build and
their results.

When a build or one of its steps fails, developers typically scavenge
the logs for information related to the source of the error.
This
manual activity is tedious and prone to
error~\cite{santolucito2018statically}.
As a single build log can be
over 50 megabytes large~\cite{beller2017oops}, finding the small chunk
of information which is linked to the actual error, is akin to
searching for a needle in a haystack.
The large numbers of unrelated, but ambiguous
warnings, which many build
logs regularly contain, further compound this problem.
In addition to helping
understand and fix build failures, the information richness of build
logs enables a plethora of other applications: with the help of build
logs, we can better understand and categorize CI failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to do ``predictive CI'' on the build or test outcome
and
duration~\cite{ni2017cost,bisong2017built,haghighatkhah2018test,
machalica2019predictive},
or investigate the role of automated static analysis tools (ASATs,
\cite{beller2016analyzing}) in CI
builds~\cite{zampetti2017open}.

@@

Despite its central role in helping developers and in enabling
novel, on-ward processing applications,
the process of how to
%automatically
% ^ we survey manual as well, though leaving it out includes too much
systematically
analyze build logs has thus far not been systematically
overviewed as a research topic, bringing us to the first
research question:
\begin{simplebox}[minipage boxed title*=-5cm]{\textbf{Research Question
1}}
What is the state-of-the-research on analyzing build logs?
\end{simplebox}

To answer this question, we start our investigation---pictured in
\Cref{fig:overview}---with an
extensive literature survey (step 1).
The survey identifies various strategies to
analyze build logs, among which are developing
custom parsers, using regular expressions, doing manual inspections,
searching for keywords, or performing natural language processing.
However, there is currently no guidance on when to use which of these
approaches.
Unfortunately, our survey also shows that
less than half of the 61 works which rely on build log analysis
describe it in sufficient detail and that
the few available implementations are all based on either a custom parser
or regular expressions.
These two approaches in particular are problematic, as their creation
and maintenance is manual and expensive, does not generalize across build
% TODO Moritz: should we say 'our' experience? or 'one authors'?
% the latter might be too cumbersome
logs from different build tools, and is, as the experience of
TravisTorrent has shown, prone to error.
In fact, some even ``strongly discourage'' the creation of such
parsers~\cite{urli2018design}.
In summary, the outcome of our literature survey shows the need for a new
generation of automated techniques to analyze build logs, which are
openly available and documented in detail, together with
guidance on when to use which technique.

@@

% turned into an itemize again cause the 'sentence' around it was
% rather broken and itemish anyways
To fill this gap, we adapted and implemented three novel, prototypical
implementations to
automatically analyze build logs (step 2).
The ideas behind the implementations are based on promising approaches
adopted from other research fields and existing build log analysis
techniques.
In this article, we present three techniques that analyze a build log
by retrieving a substring (chunk) that describes a
specific information:
\begin{itemize}
	\item \emph{Program Synthesis by Example (PBE)}, a technique
	able to
	automatically synthesize regular expressions,
	\item a technique based on information
	retrieval, using \emph{Common Text Similarity (CTS)}
	\item inspired by the manual adhoc-search for key parts in the
	log, a
	technique employing \emph{Keyword Search (KWS)}
\end{itemize}
% a technique 1) able to automatically synthesize regular expressions,
% called \emph{Program Synthesis by Example (PBE)}, a technique
% 2) based on information
% retrieval, using \emph{Common Text Similarity (CTS)}	and 3)
% inspired by the manual adhoc-search for key parts in the log, a
% technique employing \emph{Keyword Search (KWS)}.
These three approaches---synthesized regular expressions,
text similarity,
keyword search---have
different strengths and weaknesses.
As a consequence,
in the second research question we ask:

@@

To answer this question
we conduct an empirical study assessing the performance
of these chunk retrieval techniques in
real-world projects under a variety of conditions (step 3).
The study is based on the \emph{LogChunks} data
set~\cite{brandt2020logchunks}, which encompasses 797 build logs from
80 popular open-source projects and a broad range of different
build tools and programming languages.
Our results show that there is no technique that in general outperforms
the others.
PBE has the best average precision of 95\%, while KWS has the largest
recall with 70\% on average.
However, CTS shows the best F$_{1}$-score
of 51\% and all of the three techniques
clearly surpass a random baseline.

Due to this ambiguity, we develop a set of guidelines
on how to choose the most suitable build log analysis
technique for the task at hand (step 4).
We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
The results PBE produces are suited for automatic on-ward processing.
CTS is
well-suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a found keyword---makes it generally
unsuited for automatic on-ward processing.
Instead, KWS requires a human
to further inspect and interpret the output chunk.
A hybrid solution combining the three techniques might be the best
all-round solution.
Finally, we give a roadmap to guide future research in the field
of build log analysis.

In short, this article contributes
\begin{itemize}
\item the first systematic survey of the emerging field of build log
analysis.
\item three prototypical open-sourced implementations of
promising chunk retrieval techniques.
\item a large empirical study comparing the three techniques.
\item systematic guidelines on when to use which technique.
\item a detailed research map to inspire future contributions to the
field.
% (resolved) only point with no adjective
\item a documented replication package~\cite{brandt2020chunk-replication}.
\end{itemize}

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system
logs which are produced during runtime of the system.
The first difference is that build logs are produced during runtime
of the build of the system, not during runtime of the system itself.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} depicts two
exemplary lines from such a system log.

@@

\subsection{Literature Selection}
The first step in a systematic literature survey is to
establish which works one wants to cover and how to find
them~\cite{kitchenham2009systematic}.
Since the field of build log analysis is relatively young, we
surveyed as broad a population of scientific material as possible.
We did not limit the sources to specific Software Engineering
venues~\cite{petersen2015guidelines}, but included all sources
monitored by Google Scholar, including Bachelor's theses, Master's
theses, and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases because it is the search engine for scholarly material
with
the broadest and most current index, including preprints.

\Cref{fig:lit-survey} gives an overview of the literature selection
process.
An initial Google Scholar search for variations of ``build log''
returned more than 2.2 million results on the 15th of March 2020, many
of which stem
from unrelated fields such as construction or wood working.
We refined the search criteria to exclude such obvious
non-related fields and to exclude works on systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results.
% Our replication package documents the full search
% queries~\cite{brandt2020chunk-replication}.
To make handling so many search results feasible, we employed a
three-pass filtering strategy:

First, we filtered articles based on (a) their title and (b) the
contextual information Google Scholar displayed on its search results
overview page.
We included works at large that seemed to bear a resemblance to software
engineering and building software.
Following Kitchenham's protocol~\cite{kitchenham2009systematic},
we excluded
works written in a language unintelligible to the authors
(i.e., not English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

@@

Specifically, we were interested in
\begin{itemize}
  \item what information the researchers retrieve from build logs and
  what this information is used for (\textbf{RQ1.1}),
  \item which techniques they to process the build logs,
  in how much detail the techniques are described,
	and whether their implemented tools are published
  (\textbf{RQ1.2}),
  \item what kind of build logs the researchers analyze
	and the origin of these logs (\textbf{RQ1.3}).
\end{itemize}

This lead us to the following research questions:
\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What is the state-of-the-research on analyzing build logs?}
\begin{itemize}[leftmargin=1cm]
  \item[\textbf{RQ1.1:}] What information is targeted in build logs?
  \item[\textbf{RQ1.2:}] Which analysis techniques exist?
  \item[\textbf{RQ1.3:}] Which kinds of build logs are subject to
  analysis?
\end{itemize}
\end{simplebox}

To support the data extraction we created a template that we filled out
for each of the 61 works.
The template contained questions to answer RQ 1.1 through RQ 1.3,
for example ``Is the technique explained in detail?'' and
``What is the source of the build logs?''
As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping survey with a
pilot:
both authors extracted data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 56 articles evenly between the two authors
and inspected each article as closely as necessary to fill out the
template
with confidence,
starting from occurrences of ``log'' or ``build'' within the article's
text.

@@

\subsubsection{Descriptive Study Statistics}
To be done.
% TODO moritz: age of papers (graph), kind of
% publication (master, paper)
% published (y, n)

@@

\raisebox{0.8mm}{Information Retrieval} &
\raisebox{0.8mm}{
\cite{hassan2017change,lindqvist2019detection,ren2018automated}
} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\

\raisebox{0.8mm}{Analysis} &
\raisebox{0.8mm}{
\cite{sulir2016quantitative,haghighatkhah2018test,durieux2019critical}
} &
\includegraphics[width=0.55\columnwidth]{img/lit-sur/techniques-no-guidelines-cropped_3.pdf}
\\

@@

Of the 61 articles we inspected, 43 mentioned how they analyzed
build logs.
However, only 16 of these described their method in detail.
\Cref{tab:litsur:techniques} presents the techniques that the articles
described for analyzing build logs.

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''
~\cite{rahman2018impact}.
For such imprecise statements it was often not clear if they describe
a complex parser (including a tokenizer, lexer, etc.) or refer to a
simpler parsing tool based on regular expressions.
If we saw evidence for the latter, we assigned the tag
``regular expression'' instead of the tag ``parser.''
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with manual categorizing build logs and selecting
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.
Another 8\% employed machine learning such as natural language
processing or information retrieval techniques.
The rather general tags ``analysis'' and ``scan'' in
\Cref{tab:litsur:techniques}, stem---similar to the imprecise
statements about parsing---from the
words authors used to describe their techniques.
%such as
%``automatically analyzed [the build logs]''.

@@

Only 7\% of the proposed methods claimed to be language-agnostic, while
several described they are covering multiple source build tools.
The majority of articles (42\%) used logs from a source that no
second article
targeted.
In 38\% of cases, the articles collected build logs from Travis
CI~\cite{travisci2019webpage};
18\% of the articles processed logs from TravisTorrent; and 8\% included
build logs from industrial projects.
The number of logs analyzed varied greatly between the articles
and we were not always able to
extract it confidently.
23 of the articles analyzed more than a thousand build logs, while one
investigated up to 122 million logs.
Of the surveyed articles, 31\% claim to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which
Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying},
and several other articles studied.

@@

% ----- descriptive / general -----
Our literature survey shows that despite its relatively young age
(first mentioned in a paper in 2001, vast majority after
2016), \textbf{build log analysis is an established technique}
in the literature, with 61 articles making use of it.

@@

% ----- information RQ 1.1 -----
In regards to the information which was targeted, \textbf{most build log
analysis approaches aim to retrieve the reason a build failed} by
targeting ``errors'' or ``failing tests''.
% which is what we also do in LogChunks
\textbf{The majority of approaches retrieves specific chunks of
text from the logs.}

% ----- technique RQ 1.2 -----
\textbf{Most log analysis is automated.} Automated approaches are
necessary as many studies target a large number of builds.
Articles
which explicitly mention a manual analysis of the logs, mostly
restrict their work to few logs or take a ``representative sample to
make manual analysis feasible''~\cite{zolfagharinia2017not}.
We observed \textbf{very little reuse of build log analysis tools},
even though
\textbf{developing automated analysis tools requires a lot of effort}.
Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone,'' while other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.
% not sure if we can state that (no data on this, more 'impression')
In addition, we rarely saw articles justifying why they use a certain
technique to analyze build logs as there is a general \emph{lack of
guidance on when to use which technique}.

% ----- kinds of build logs RQ1.3 -----
Several articles noted the variety in log formats for different build
tools, which requires customizing build log techniques for
every supported build tool.
% The following two sentences are passive, however flipped around
% have a different meaning (we want build tool -> study count)
\Cref{fig:litsur:log_producer} shows that only a few distinct
build tools are the target of
many studies and tools (roughly, area A).
The majority of build tools was however only targeted by one
article each (roughly, area B).
From this we follow, that the large tail of
\textbf{seldomly-investigated
environments would benefit from the
existence of a more generic solution.}
In most cases, \textbf{researchers analyze build logs from open
source projects}, obtained through Travis CI.
In addition, we saw several articles that targeted the logs of only
one specific software project.

% ----- custom build tools -----

% What about the alternate solution, installing plugins to report
% precise data? Ie., maven plugins
% - Lots of effort
% - not a generic solution
% - how to analyze history where such plugins were not installed
% - unclear which information you want

% TODO Moritz: is this okay (here)?
% maybe it's better to add it to Threats to Validity with the heading:
% \texbf{Modifying the Build Process.}
An alternative to analyzing build logs, would be to integrate plugins
that report precise data about the build process, such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin}.
However, such plugins are not available for all build tools and they
do not enable the analysis of historical data---builds that ran before
the plugin was installed cannot be analyzed.
In addition to that, when using such a plugin, one would have to know
beforehand which information should be reported from a build, whereas
build log analysis enables us to analyze the wide range of information
already baked into the build logs.
Their advantage is that they can precisely report the information which
is required.
A downside of plugins, which are integrated into the CI flow similar to
Deflaker by Bell et al.~\cite{bell2018deflaker}, is that they are unable
to work on historic
builds, i.e., before the moment they were being deployed and that if
information has not been recorded from the beginning but turns out to
be important, it is impossible to retrieve later.

% \textbf{modifying the build output is not helpful}, as well as returning
% precise data through other documents than the log.
% This requires access to the build configuration, is also specific to
% the used build tool and prohibits the analysis of
% historical data.

% \textbf{ embed custom plugin discussion}
% %  very disconnected here
% From this, we follow: 1) there is a niche for custom parsers or even
% specialized plugins integrated into the build process to make
% post-factum
% parsing of logs redundant.
% Plugins  similar to, for example, the Maven plugin to report test
% executions Bell et al.
% % cite
% wrote, have the disadvantage of not being able to work on historic
% builds.
% 2) the large tail of seldomly-investigated environments would benefit
% from the existence of a more generic
% solution.

% We implement build tool agnostic build log analysis
% techniques, which are customized to a specific project by
% the training examples which the user provides.
% In addition, the \emph{LogChunks}~\cite{brandt2020logchunks}
% data set we chose for our evaluation covers a broad range of build
% environments.
% % Further, the training examples also specify the log chunk targeted
% by the
% % retrieval, which enable our chunk retrieval techniques to retrieve
% % a broad range of information from build logs.

@@

% Alt:
%\section{Analyzing Build Logs with Chunk Retrieval Techniques}
\section{Chunk Retrieval Techniques}
\label{sec:techniques}
%TODO: really? currently the discussion does not show this 'need'!
The outcome of our literature survey shows the need for a new
generation of techniques to analyze build logs.
In this section we argue how our research sets the first step to
address this need by introducing three techniques that retrieve
specific substrings (chunks) from build logs.
We present defining characteristics of these so called \emph{chunk
retrieval techniques} and present the techniques we developed in
detail.
% TODO At the end of what?
At the end we illustrate how the techniques operate on a concrete
example and sketch out their implementation.

@@

\subsection{A New Generation of Build Log Analysis Techniques}
In summary, the results of the literature survey show the need for
build log
analysis techniques that are \emph{automated},
\emph{agnostic of the build environment}
and can be \emph{configured with little effort}.

% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
\emph{Examples consist of logs and corresponding chunks.}
The user provides these examples and keywords, which are easy to adapt
whenever the user adds new examples.
A user---a developer or a researcher who wants to analyze
build logs---provides these examples and keywords, which can simply be
